Bringing the Power of eBPF to Open vSwitch

### 2018-10-29 paper.tex
Minor modification in ch 4

\section{Userspace OVS with AF\_XDP}
\label{sec:af_xdp}
\begin{figure}
{\scriptsize
\begin{verbatim}
              |   +-------------------+
              |   |    ovs-vswitchd   |<-->ovsdb-server
              |   +-------------------+
              |   |      ofproto      |<-->OpenFlow controllers
              |   +--------+-+--------+ 
              |   | netdev | |ofproto-|
    userspace |   +--------+ |  dpif  |
              |   | AF_XDP | +--------+
              |   | netdev | |  dpif  |
              |   +---||---+ +--------+
              |       ||     |  dpif- |
              |       ||     | netdev |
              |_      ||     +--------+  
                      ||         
               _  +---||-----+--------+
              |   | XDP program +     |
 Linux kernel |   |   eBPF map        |
              |_  +--------||---------+
                           ||
                        physical
                           NIC
\end{verbatim}
}
\vspace{-1.0em}
\caption{OVS Architecture AF\_XDP}
\label{ovsafxdp}
\vspace{-1.0em}
\end{figure}

\subsection{Datapath and Netdev Interface}
OVS provides a userspace datapath interface (dpif) implementation, called
dpif-netdev. The dpif-netdev userspace datapath receives and transmits
packets from its userspace interface.  One major use case of dpif-netdev
is OVS-DPDK, where the packet reception and transmission
are all conducted in DPDK's userspace
library. The dpif-netdev is designed to be agnostic to how the network
device accessing the packets, by an abstraction layer called netdev.
Therefore, packets can come from DPDK packet I/O library,
a Linux AF\_PACKET socket API, or in our case, the AF\_XDP socket interface,
as long as each mechanism implements its own netdev interface.
Once dpif-netdev receives a packet, it follows the same mechanism performing
parse, lookup the flow table, and apply actions to the packet.

@@

\subsection{umem memory management}
In order to properly program Rx/Tx/Fill/Completion rings, we implement a
umem memory management layer, call umempool. Basically, it's a data structure
maintaining the unused/available elements in umem with GET and PUT access
APIs. We will demonstrate the use case of umempool with packet reception and
transmission in the following subsections.

@@

Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the Fill ring and
the Rx ring for receiving packets from XSK.  For simplicity, in this example,
we assume that there are only eight umem buffers, and each buffer's
size is 2KB.  Initially, at step 1, ovs-vswitchd pushes four available umem
elements into
the Fill ring and waits for incoming packets.  When there are incoming packets,
at step 2, the Fill ring's four buffer elements have been consumed and moved
to the Rx ring.  In order to keep receiving packets, ovs-vswitchd gets another
four available umem elements from the umempool, and fills into Fill ring (step 3).
Then, ovs-vswitchd creates the metadata needed for the four packet
buffer \{1, 2, 3, 4\}, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 
the dpif-netdev layer for parse, lookup and action executions.
Finally, when ovs-vswitchd finishes processing the umem buffer, a recycle
mechanism is triggered to place this buffer back to umempool (step 5).
Step 5 makes sure that there are always available elements in Fill ring, so
that the underlying XDP program in kernel can keep processing packets 
while the userspace ovs-vswitchd is processing the previous
received packets on Rx ring.
When step 5 finishes, ovs-vswitchd goes back to step 2, waiting for
new packets.

@@

\subsubsection{PMD netdev}
In the beginning, our initial prototype of OVS-AFXDP only provides $0.5Mpps$
throughput for rxdrop.
We observe that when dropping/forwarding packets under OVS-AFXDP, two processes
below can easily reach 100\% CPU utilization:
\begin{itemize}
\item ovs-vswitchd: This is the process that keeps doing the send and receive steps
in Figure~\ref{afxdprx} and~\ref{afxdptx}.
\item ksoftirqd/core\_id: This is the kernel software interrupt thread handling the
incoming packets, triggering XDP program to pass packets to the XSK, and also
processing transmission.
\end{itemize}
\mycomment{Compared to xdpsock, ovs-vswitchd adds on top of it a umempool,
and packets parsing, lookup and action execution.}
With further investigation using Linux perf tools, our first improvement is to
to enable OVS's Poll-Mode-Driver (PMD) mode to the afxdp netdev.

@@

With the above design change, we find that using the LIFO-ptr\_array in
Section~\ref{memorym} with separated packet metadata allocation yields the
best performance. It is because both data structures have better spatial
locality and there are more batching friendly.
For example, accessing 32 packet metadata in an array incurs less cache
misses than accessing the packet metadata 32 umem chunks.
The ptr\_array for umempool management outperforms the other two designs
for the similar reason. With the above design decisions, the OVS rxdrop
can achieve similar performance as the baseline xdpsock rxdrop at around
$19Mpps$.

\subsubsection{Batching Send Syscall}
%574.790 M/sec  
With all the aforementioned optimizations, We continue measuring the
performance of OVS-AFXDP l2fwd and observe only $4Mpps$, compared to
$17Mpps$ baseline xdpsock-l2fwd.
We find that the OVS PMD thread under rxdrop has much fewer context
switches compared to the l2fwd, indicating that
the PMD process spends much more time in kernel space than in userspace.
By using strace, we find that the OVS-AFXDP l2fwd experiment calls sendto
system call at very high rate. It is because we design to check the
completion of send immediately after issusing send as shown in step 3 and
4 in Figure~\ref{afxdptx}.
We modify the design by calling sendmsg syscall (step 3) only when TX ring is
close to full, e.g., when 3/4 ring elements have been used.
For example, instead of issuing sendmsg syscall for a batch of 32 packets
and making sure they are finished, we only issue send when there are 512
outstanding packets.  With this change, the OVS-AFXDP l2fwd experiment
can achieve around $14Mpps$.

>>>>

\section{Userspace OVS with AF\_XDP}
\label{sec:af_xdp}
\begin{figure}
{\scriptsize
\begin{verbatim}
              |   +-------------------+
              |   |    ovs-vswitchd   |<-->ovsdb-server
              |   +-------------------+
              |   |      ofproto      |<-->OpenFlow controllers
              |   +--------+-+--------+ 
              |   | netdev | |ofproto-|
    userspace |   +--------+ |  dpif  |
              |   | AF_XDP | +--------+
              |   | netdev | |  dpif  |
              |   +---||---+ +--------+
              |       ||     |  dpif- |
              |       ||     | netdev |
              |_      ||     +--------+  
                      ||         
               _  +---||-----+--------+
              |   | XDP program +     |
 Linux kernel |   |   eBPF map        |
              |_  +--------||---------+
                           ||
                        physical
                           NIC
\end{verbatim}
}
\vspace{-1.0em}
\caption{OVS Architecture with AF\_XDP}
\label{ovsafxdp}
\vspace{-1.0em}
\end{figure}

\subsection{Datapath and Netdev Interface}
OVS provides a userspace datapath interface (dpif) implementation, called
dpif-netdev. The dpif-netdev userspace datapath receives and transmits
packets from its userspace interface.  One major use case of dpif-netdev
is OVS-DPDK, where the packet reception and transmission
are all conducted in DPDK's userspace
library. The dpif-netdev is designed to be agnostic to how the network
device accessing the packets, by an abstraction layer called netdev.
Therefore, packets can come from DPDK packet I/O library,
a Linux AF\_PACKET socket API, or AF\_XDP socket interface,
as long as each mechanism implements its own netdev interface.
Once dpif-netdev receives a packet, it follows the same mechanism performing
parse, lookup the flow table, and apply actions to the packet.

@@

\subsection{umem memory management}
In order to properly program Rx/Tx/Fill/Completion rings, we implement a
umem memory management layer, call umempool. It is a data structure
maintaining the unused/available elements in umem with GET and PUT access
APIs. We will demonstrate the use case of umempool with packet reception and
transmission in the following subsections.

@@

Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the Fill ring and
the Rx ring for receiving packets from XSK.  For simplicity, in this example,
we assume that there are only eight umem buffers, and each buffer's
size is 2KB.  Initially, at step 1, ovs-vswitchd pushes four available umem
elements into
the Fill ring and waits for incoming packets.  When there are incoming packets,
at step 2, the Fill ring's four buffer elements will be consumed and moved
to the Rx ring.  In order to keep receiving packets, ovs-vswitchd gets another
four available umem elements from the umempool, and fills into Fill ring (step 3).
Then, ovs-vswitchd creates the metadata needed for the four packet
buffer \{1, 2, 3, 4\}, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 
the dpif-netdev layer for parse, lookup and action executions.
Finally, when ovs-vswitchd finishes processing the umem buffer, a recycle
mechanism is triggered to place this buffer back to umempool (step 5).
Step 5 makes sure that there are always available elements in Fill ring, so
that the underlying XDP program in kernel can keep processing packets 
while the userspace ovs-vswitchd is processing the previous
received packets on Rx ring.
When step 5 finishes, ovs-vswitchd goes back to step 2, waiting for
new packets.

@@

\subsubsection{PMD netdev}
In the beginning, our initial prototype of OVS-AFXDP only provides $0.5Mpps$
throughput for rxdrop.
We observe that when dropping/forwarding packets under OVS-AFXDP, two processes
below can easily reach 100\% CPU utilization:
\begin{itemize}
\item ovs-vswitchd: This is the process that keeps doing the send and receive steps
in Figure~\ref{afxdprx} and~\ref{afxdptx}.
\item ksoftirqd/core\_id: This is the kernel software interrupt thread handling the
incoming packets, triggering XDP program to pass packets to the XSK, and also
processing transmission.
\end{itemize}
\mycomment{Compared to xdpsock, ovs-vswitchd adds on top of it a umempool,
and packets parsing, lookup and action execution.}
With further investigation using Linux perf tools, our first improvement is
to enable OVS's Poll-Mode-Driver (PMD) mode to the AF\_XDP netdev.

@@

With the above design change, we find that using the LIFO-ptr\_array in
Section~\ref{memorym} with separated packet metadata allocation for
metadata allocation yields the
best performance. It is because both data structures have better spatial
locality and they are more batching friendly.
For example, accessing 32 packet metadata in an array incurs less cache
misses than accessing the packet metadata in 32 umem chunks.
The ptr\_array for umempool management outperforms the other two designs
for the similar reason. With the above design decisions, the OVS rxdrop
can achieve similar performance as the baseline xdpsock rxdrop at around
$19Mpps$.

\subsubsection{Batching Send Syscall}
%574.790 M/sec  
With all the aforementioned optimizations, We continue measuring the
performance of OVS-AFXDP l2fwd and observe only $4Mpps$, compared to
$17Mpps$ baseline xdpsock-l2fwd.
We find that the OVS PMD thread under rxdrop has much fewer context
switches compared to the l2fwd, indicating that
the PMD process spends much more time in kernel space than in userspace.
By using strace, we find that the OVS-AFXDP l2fwd experiment calls sendto
system call at very high rate. It is because we design to check the
completion of send immediately after issusing send as shown in step 3 and
4 in Figure~\ref{afxdptx}.
We modify this design by calling sendmsg syscall (step 3) only when TX ring is
close to full, e.g., when 3/4 ring elements have been used.
For example, instead of issuing sendmsg syscall for a batch of 32 packets
and making sure they are finished, we only issue send when there are 512
outstanding packets.  With this change, the OVS-AFXDP l2fwd experiment
can achieve around $14Mpps$.
